Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2022-09-09 ba_resubmit/response_to_reviewers.tex

		\pointRaised{E}{%
	The loss function on p. 9 is out of Sadinle’s paper and it inherits the typo therein $(\theta_{11}, \text{if} \; Z_j, \;\hat{Z_j}, \ldots)$.
	}

	\reply{
	\textbf{The previous expression $Z_j, \hat{Z}_j \leq n_1$ was meant to be understood as $Z_j\leq n_1, \hat{Z}_j \leq n_1.$ We agree that the latter expression is more clear. This has been updated in our revision.}
	}

		\pointRaised{E}{%
	In the statement of Lemma 1 on p.13, B comes out of nowhere.  (Thinking I had missed something, I went back to the previous pages only to find out that B would be then defined in the proof.)
	}

	\reply{
	\textbf{We have updated the Lemma to include the definition of $B$ as well as other variables in the Lemma.}
	}

@@

	\reply{
		Punctuation for all math has been checked.
	}

\pointRaised{E}{%
	Confusion can also arise from lack of clarity in the exposition.  Take comment 4 from the AE, for example.  Whose marriage certificates and whose birth records are involved?  What records are in $X_1$ and what records are in $X_2$?  The reader should not be expected to go back to Newcombe’s paper to clarify the issue.
	}

	\reply{%
\textbf{We agree with this comment, and we have removed this example as it does not add to the paper. Furthermore, we have gone through the paper to improve the overall exposition and clarity of the writing.}
}

\pointRaised{E}{%
	Other comments that were raised are as follows: This is not an exhaustive list, but only a set of examples meant to point out what makes the paper hard to read and how it can be improved.  We summarize these below. 
}

\reply{	
	\begin{enumerate}
	\item p. 12, second paragraph:  insert “is” after Section 3.1.  \textbf{This has been fixed.}
	\item p. 15:  the first sentence of the last full paragraph is either missing a verb (“we observe”?) or it includes an extra “that.” \textbf{We are not able to find this error. Go check this.}
	\item As I said, I do not believe that we, as reviewers, were able to uncover all the problems, and I strongly encourage the authors to do their part, as they should, to improve the presentation and eliminate all typos and inconsistencies. \textbf{This has been done by all authors.} \todo
	\item The AE finds the simulation studies to be incomplete and gives detailed suggestions on how those should be improved.  The AE also gives important suggestions about other aspects of the manuscript that must be carefully considered and addressed.  All of the AE’s comments are right on the mark. \textbf{We have provided detailed comments to the AE.}
	\end{enumerate}
	}

%	\reply{%
%	\begin{enumerate}
%		\item This change has been made
%		
%		\item \textcolor{red}{I could not find this error}
%		
%		\item We have read the paper for typos and inconsistencies, and thank the reviewers for bringing these to our attention. 
%		
%		\item We have responded to the AE's comments about the simulation studies
%	\end{enumerate}
%}

@@

	\setcounter{responsectr}{0}

	\pointRaised{R1}{%
	This manuscript addresses the bipartite matching problem in data linkage and proposes a new fast computation version of “Beta Linkage” to allow such matching to take place for pairs of large databases without resorting to ad hoc blocking. It is built upon the Felegi-Sunter model of record linkages, where linkages are made within strata of agreement between records across the two databases, and relies heavily on advances made by Sadinle (2017). Overall, the paper presents a practical implementation of a modified Beta Linkage approach to	matching. It demonstrates that even though the new method suffers from the limitations of not seamlessly enforcing one-to-one matching, it is feasible for large datasets with modest numbers of features on which to match. Though the novelty of the presented methodology is primarily in the computational approach, I believe it to be a real advance for practical implementation of record linkage in modern applications.

	I have no major concerns about the content or accuracy of the presentation. However, substantial improvements are warranted to confirm my impression. In addition to the recommendation made by Reviewer 1 to include more clarity of the novel contributions of this manuscript, I believe that the explanation of the method should be written more clearly, the simulation comparisons should be repeated/expanded, and various features of the simplification should be fleshed out a bit more. Finally, more practical advice would be welcome. All of these points are more specifically described in the itemized list below
	}

	\reply{%
	\textbf{We thank the reviewer (AE) for noticing our computational novelties that expand upon the prior limitations of the prior work of Sadinle (2017). Our review paper clarifies our novel contributions, improves upon the exposition of the writing, provides a practical guidance for users, and expands the simulations. We expand on these points further below.} \todo
	}

@@

	\pointRaised{R1}{%
		Overall, the notation is, somewhat by necessity, quite extensive. However, I’m not convinced that it’s entirely consistent. A glossary in the appendix may help the reader navigate the paper more easily.}
\reply{%
	\textcolor{blue}{I think this is an excellent idea and it could make sure that we don't miss anything. It would really help the reader. You could make a table of the notation, and Neil does this quite well in the Marchant 2021 paper as an example.} \todo
}

\pointRaised{R1}{%
	Section 2, line 1: It may help to explicitly define X1 and X2 as being vectors of indices as implied by equations (1) and (2). 
}
\reply{
	We thank the reviewer for the suggestion, which has been incorporated into the paper in the following way: ``Consider two databases $X_1$ and $X_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively. Without loss of generality, denote files such that $n_1 \geq n_2$. We follow the convention established by Sadinle (2017) and say ``record i $\in X_1$" rather than the more compact $x_{1i}$ in order to avoid double subscripts."

@@

%	\textcolor{red}{OPTION 2: We agree that the notation provided conflates the records themselves with the record indices. However, this notation was established in Sadinle 2017 and has been replicated in further work (Wortman 2019, Alishen-Guendel, one of Jerry's students), and we have decided to use it as well for consistency within the literature.}
%	\textcolor{blue}{We appreciate the suggestion, however, we believe it would be confusing to the existing prior work by Fellegi and Sunter (1969) that has been established as well as others that have extended the work, such as Sadinle (2014+), Wortman (2019), and Asheshin-Gundel (2020+). Given that it's widely used in this literature, we continue to utilize it.}
}

\pointRaised{R1}{%
Section 2.2, beta distribution: the $n_{12}(.)$ notation needs to be defined.
}
\reply{%
	\textbf{This has been defined.}
}

\pointRaised{R1}{%
Section 3, 4th full paragraph: My understanding of the example of matching birth records and
marriage certificates (presuming monogamy) may not quite be consistent with the algorithm as
presented, as I would think that one certificate (element in X1) would match multiple birth
records (elements in X2) which violates the size ordering. Can the algorithm be easily adjusted
for this reverse sizing?
}

\reply{
	\textbf{Thank you for the interesting comment/question. In this paper, our main goal and applications deal with bipartite record linkage (such that we can make fair comparison with prior work), which is consistent with the algorithm that we present in the paper. We have removed the confusing text mentioned by the reviewer. We are working on extensions beyond bipartite record linkage in other work, where the reverse sizing is not an issue, however, this is beyond the scope of this paper.}
}

@@

%\textcolor{blue}{This is a good point. I wonder if we could perhaps remove a lot of the notation by Sadinle/FS and redo the notation such that the ordering wouldn't matter. Also, it would make more sense to follow with the examples that we use in the paper -- NLTCS and EL Salvador so that we don't confuse the reader. Let's talk through this point as this would be a big change. If the algorithm and notation can be easily modified to reverse things, this would be a nice addition to the paper and improvement over Sadinle.}

\pointRaised{R1}{%
5. Section 3, full model: $I_obs(.)$ seems to be taken from a different paper dealing with missing fields, which is not at all discussed in this manuscript. This consideration of missing fields and the accompanying discussion of missing data mechanism should be included if such data is to be considered for this manuscript. This comment also applies to the $Gamma^{obs}$ notation in the following subsection.
}

\reply{%
	\textbf{Our missingness assumptions and the definition of $I_{obs}()$ are near the end of Section 3.}
}

@@

\pointRaised{R1}{%
	Section 3, full model: The square brackets seem to be misplaced, as the exponent includes indices of the sum contained in the brackets.
}

\reply{%
	\textbf{The expression is now corrected.}
}

\pointRaised{R1}{%
Section 3.1, I found the motivational descriptions in this section to be somewhat confusing, in that they seem to mix the posterior behavior (e.g., parameters are updated though standard multinomial-Dirichlet…) and model construction (as a function of pi, Z is a series of successes…). This subsection might be easier to understand if its purpose was more bluntly described in an opening sentence or paragraph and streamlined to focus only on the full conditional distributions for a Gibbs Sampler. In addition, the mixed motivation may have led to errors in intermediate conclusions. In particular, the derivation of the probability of $\Gamma_{.j}$ seems to be missing a factor of $u_{fl}^I(gamma_{ij}^f=l)$. This is not important to the eventual full conditional of Z, but does not seem to be correct for the distribution of Gamma as claimed.
}

\reply{%
	We have revised our presentation of the Gibbs sampler with this feedback in mind. The main body of the text now presents the final full conditionals, with revised derivations moved now in the appendix. In particular, we have included an extra step in the derivation of $p(\Gamma_{.j})$ that shows exactly how we divide by $\prod_{i = 1}^{n_1}\prod_{f = 1}^{F} \prod_{\ell = 1}^{L_f}  u_{fl}^{I(\gamma_{ij}^f=l)}$ and use proportionality to arrive that the distribution presented.
}

%\reply{%
%\textcolor{red}{\todo We should meet specifically to talk about how to best present the Gibbs Sampler. I think it be good to move as much of it as possible (at least the messy math and derivations) into the appendix.} \textcolor{blue}{I agree that these should be moved to the appendix as well.}
%
%We have included an extra step in the derivation of $p(\Gamma_{.j})$ that shows exactly how we divide by $\prod_{i = 1}^{n_1}\prod_{f = 1}^{F} \prod_{\ell = 1}^{L_f}  u_{fl}^{I(\gamma_{ij}^f=l)}$ and use proportionality to arrive that the distribution presented. \textcolor{red}{I want you guys to look at it though. Its pretty ugly.}
%}

@@

\pointRaised{R1}{%
Section 3.1, first equation on page 8: There is an extra parenthesis in the exponent of (1-pi).
}
\reply{%
	\textbf{This has been corrected.}
}

\pointRaised{R1}{%
Section 3.1, pmf for $Gamma_{.j}$ and full conditional for Z: The notation in the final equation may be more clear if you replace i with $z_j$, as in $w_{z_j j}$. I believe that “$n_1+1$” should be “$n_1+ j$” in several places.
}

\reply{%
	\textbf{We have changed $w_{ij}$ to $w_{z_j, j}$ in the pmf for $\Gamma_{.j}$ and the full conditional for $Z_j$. We have been sure that “$n_1+1$” has been changed to “$n_1+ j$” throughout the paper.}
}

\pointRaised{R1}{%
	Section 3.1, last equations on page 8: There seems to be some lack of specificity in describing these equations as full conditional distributions vs. steps in the Gibbs Sampler algorithm. It would be helpful to provide a brief justification of integrating out pi.
}
\reply{%
	\textbf{We have revised the presentation of the Gibbs sampler based on this feedback. The main body of the paper presents the results, and revised derivations (including justification for integrating out $\pi$) have been placed in the appendix.}
}

\pointRaised{R1}{%
	Section 4.2, SEI procedure. I’m not sure I follow the description of the SEI procedure. (1) My understanding of SEI is that for patterns with lots of possible pairs, no single pair is likely to (should) be identified in the posterior distribution. (2) Thus, you take a small subsample of the records of size $S < H_{j_p}$, and store these in $R^{SEI}$ rather than the complete index of pairs in each R. (3) Can you quantify the computational savings due to this method? (4) Do you have advice for how small S can be? (5) In your simulations, how does the use of this method or choice of S affect the results?
}

\reply{%
	We have numbered each sentence in the comment above to clearly address each point. (1) and (2) are correct understandings of the SEI procedure. For (3), we quantify computational savings of SEI in Section 4.2, where we state:
	"Rather than storing \(n_1 \times n_2\) record labels, SEI allows us to store at most \(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be." SEI does not affect the computation time of the Gibbs sampler, since the complexity of the sampler is $O(n_2 P)$, regardless of how many comparison vectors are of pattern $P$.

	For (4), we have included the following sentence in the paper ``In practice, we recommend $S=10$, as this reduces the number of stored indices for highly unlikely record pairings, but does is not likely to eliminate any of the indices for likely matches."

	For (5), we have included the following sentence in the paper ``Choosing $S$ too low, like $S=1$ or $S=2$, can concentrate undue mass on unlikely matches and distort linkage results." 
}

\pointRaised{R1}{%
Section 4.4, Assumptions and definitions should be included in the statement of this lemma.
}

@@

\reply{%
	\textbf{The lemma now includes assumptions/definitions for all variables used within the statement.}
}

\pointRaised{R1}{%
Section 5.1 simulations – there is a detailed comparison for one set of simulation settings, including a binary definition of a “match”. Do these results differ as these settings change?
}

\reply{%
	Regardless of the specific construction of the comparison vectors, the computational complexity the Gibbs sampler under BRL is $O(n_1 n_2)$, where as it is $O(n_2 P)$ for fabl. The exact shape of the lines will be different, but in all cases, fabl removes dependence on $n_1$ in the computational complexity. To illustrate this, we have included in the appendix an additional simulation study under different comparison vector settings.
}

%\reply{%
%	Regardless of the specific construction of the comparison vectors, the computational complexity the Gibbs sampler under BRL is $O(n_1 n_2)$, where as it is $O(n_2 P)$ for fabl. The exact shape of the lines will be different, but in all cases, fabl removes dependence on $n_1$ in the computational complexity. 
%	
%	\textcolor{red}{I think this is clear in the paper, so I did not add anything to the actual text.} \textcolor{blue}{I'm not sure that you answered the reviewers question. You show a comparison for one set of simulation settings, but if you change the settings what would occur? I think the reviewer is asking for more clarify regarding how the lines would change. You may wish to give a few different scenarios to illustrate how things change?}
%}

@@

\pointRaised{R1}{%
	Section 5.2 and 5.3 simulations – are these a single simulated datasets? Do these same results hold over multiple simulated datasets, Or was this just a chance result for each setting?
}

\reply{%
	In the caption under Figure 3, we explain that we have 100 pairs datasets for each level of overlap and error. Thus the table represents results from 900 pairs of datasets. We have added this description in the body of the paper as well for clarity. 
}

@@

\pointRaised{R1}{%
Section 5.3 simulation – RR results are not presented.
}

\reply{%
	We have updated the Figure 4 with these results. Since its is presented along side NPV and PPV, for which values close to 1 indicate strong performance, we have presented decision rate (DR) results.
}

@@

	\section*{Reviewer 3}
	\setcounter{responsectr}{0}

	The paper presents a computational variant of the method introduced by Sadinle (2017). Simulation results are presented and the variant is applied on a new, significantly larger data set. A significant part of the paper is a reproduction of the paper of Sadinle. The variant is introduced in section 4. It is simple but consequential. The simulations and applications show how the variant accelerates the computation relative to the original method and reveals the trade-offs.

	The value of the paper resides in that computational complexity continues to be a difficult obstacle in the application of Bayesian statistics. The proposals of the paper enable the application of Sadinle’s approach to an extent not possible before. It is important to present the research for what it is: An extension of the method of Sadinle motivated by computational considerations. Here Sadinle supersedes Fellegi-Sunter in the sense while the starting point of Sadinle’s research is Fellegi-sunter, the starting point of this paper is Sadinle’s work. This should be made clearer, beginning in the abstract and reiterated in the conclusion.

	\reply{%
		\textcolor{red}{BEKA/JERRY: I think this reviewer wants us to explicitly mention Sadinle in the abstract. I think I had that originally, but we removed it. Please advise on the best way to address this comment.}
	}

### 2022-09-09 ba_resubmit/kundinger_resubmit.tex
updated speed simulations

\begin{abstract}
Recently, researchers have developed Bayesian versions of the Fellegi Sunter model for record linkage. These have the crucial advantage of quantifying uncertainty from imperfect linkages.  However, current implementations of Bayesian Fellegi Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks.  We propose and investigate a variation on Bayesian Fellegi Sunter models that we call fast beta linkage, or \texttt{fabl}.  Specifically, in \texttt{fabl} we propose independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead. This also allows us to complete pairwise record comparisons over large datasets through parallel computing and reduce memory costs through a new technique called storage efficient indexing. Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy.
\end{abstract}

@@

In what follows, Section~\ref{sec:review-of_prior-work} reviews the work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}. Section~\ref{sec:fast-beta-linkage}, proposes the \texttt{fabl} model, derives the Gibbs sampler for posterior inference, and provides the loss function used to calculate the Bayes estimate for the bipartite matching. Section~\ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculatoins and the scale of linkage tasks amenable to \texttt{fabl}. Sections~\ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and case studies of homicides from the El Salvadoran Civil War and the National Long Term Care Study.

@@

Importantly, this independence relaxes the one-to-one requirement from \texttt{BRL}; the Gibbs sampler under \texttt{fabl} does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\). To resolve many-to-one matchings and obtain a one-to-one estimate, we use the post processing procedure shown in Section~\ref{bayes-estimate}.

@@

We initialize $\Phi$ from random draws from the prior distributions of $\bm{m}$ and $\bm{u}$, and initialize $\bm{Z}$ to reflect no matches across datasets, that is $\bm{Z} = (n_1 + 1, \ldots, n_1 + n_2)$. We then iterativly sample $\Phi$ and $\bm{Z}$ from their respective full conditionals:

$$\bm{m}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}), \ldots, \alpha_{fL_f}(\bm{Z})),$$
$$\bm{u}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}), \ldots, \beta_{fL_f}(\bm{Z})),$$
$$p\left(Z_j^{(s+1)}  = i| \Phi, \bm{Z^{(s)}}\right) \propto
\begin{cases} 
	w_{z_j, j},  & z_j \leq n_1; \\
	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}}, & z_j  = n_1 + j, \\
\end{cases},$$
where 
$$w_{ij} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.$$
Derivations for these full conditionals can be found in Appendix \ref{app:derivations}.

@@

To increase the speed of the Gibbs sampler and expand the scale of the linkage tasks we can undertake, we propose hashing methods similar to those used by \cite{enamorado2019using} in the creation of \texttt{fastlink}, a fast and scalable implementation of the Fellegi Sunter method. The key insight is to recognize that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector. To make this more precise, let $h_1, \ldots, h_P$ denote the unique agreement patterns, and collect these unique patterns in the set $\mathcal{P} = \{h_1, \ldots, h_P\}$. Here,  $P = |\mathcal{P}|$ is the total number of unique agreement patterns. \(P\) is bounded above by \(\prod_{f=1}^F L_f\) which does not depend on \(n_1\) or \(n_2\). In this context, the integers $\{1, \ldots, P\}$ serve as hashed values that encode the same information as the orignial vectors themselves. Whenever possible, we conduct calculations over these \(P\) agreement patterns, instead of the typical \(n_1 \times n_2\) record pairs. Additionally, we store ``one-hot encodings'' of these patterns rather than the original Fellegi Sunter comparison vectors, to aid in vectorized computations. See Appendix ~\ref{app:ohe} for details.

@@

With SEI, however, we can reduce the comparison space, and its associated storage costs, while avoiding these drawbacks. Observe that all records $i \in \bm{X}_1$ that share the same agreement pattern with $j \in \bm{X}_2$ have the same Fellegi Sunter weight $w_{ij}$, and therefore the same probability when sampling \(Z_j\). Thus, we know that records $i \in r_{j_p}$  such that $H_{j_p}$ is large are very unlikely to be sampled consistently enough to be deemed a match through the Bayes estimate. We know this regardless of the form of the agreement pattern itself, or its associated probabilities. Therefore, rather than store all of these unlikely record labels, we choose to store only a small number \(S\) of them in a new nested listed $\mathcal{R}^{\text{SEI}}$. Rather than storing \(n_1 \times n_2\) record labels, SEI allows us to store at most
\(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be. \textcolor{red}{In practice, we recommend $S=10$, as this reduces the number of stored indices for highly unlikely record pairings, but does is not likely to eliminate any of the indices for likely matches. Choosing $S$ too low can concentrate undue mass on unlikely matches and distort linkage results.}Posterior calculations still attribute the appropriate weight to all records through the summary statistics in $\mathcal{H}$, and thus we can proceed with posterior inference through the memory reduced $\tilde{\Gamma}^{\text{SEI}} = \{\mathcal{P}, \mathcal{R}^{\text{SEI}}, \mathcal{H}\}$. For simplicity, we suppress ``SEI'' in future notation, and use $\tilde{\Gamma}$ without loss of generality.

@@

Instead, we use one-hot encodings of the agreement patterns $\mathcal{P}$ for more efficient calculations. Denote $H_p^m = \sum_{j=1}^{n_2} I\left((Z_j, j) \in h_p\right)$ to be the number of matching record pairs with agreement pattern $p$. The number of non-matching record pairs with agreement pattern $p$ is \(H_p^u = H_p - H_p^m\). Then, if $\bm{\alpha_0}$ and $\bm{\beta_0}$ are vectors of prior parameters for the $\bm{m}$ and $\bm{u}$ distributions respectively, the posterior update becomes

$$\bm{\alpha(Z)} = \bm{\alpha_0} + \sum_{p=1}^P H_p^m \times h_p$$
$$\bm{\beta(Z)} = \bm{\beta_0} + \sum_{p=1}^P H_p^u \times h_p.$$
These constitute $\tilde{L}$ summations over $P$ quantities, where $\tilde{L} = \sum_{f=1}^F L_f$ is the total number of levels used across all features in the linkage task.

Although sampling \(Z_j\) from a the full conditional provided in Section~\ref{gibbs_sampling} is conceptually straightforward, it becomes computationally burdensome when $n_1$ is large. This is because sampling a value from $n_1$ many options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_1$. To speed up computation, we break this sampling step into two. First, we calculate the Fellegi Sunter weight \(w_{p}\) associated with each unique pattern and sample the agreement pattern between \(j\) and its potential link. Second, we sample the record label uniformly among records associated with that agreement pattern for that particular \(j\in B\). More concretely, define \(h(Z_j)\) to be the agreement pattern between \(j\) and its potential match, and say \(h(Z_j) = h_{P+1}\) when \(Z_j = n_1 + j\). Then,

%$$P\left(h\left(Z_j^{(s+1)}\right) = p \mid \Phi, \bm{Z^{(s)}}\right) \propto
%\begin{cases} 
%	w_{p}\times H_{j_p}  & p \leq P; \\
%	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\lambda}}{n_{12}(\bm{Z}) + \alpha_{\lambda}} &   p = P + 1 \\
%\end{cases}$$
%$$P\left(Z_j^{(s+1)} = i \mid h\left(Z_j^{(s+1)}\right) = p\right) = \begin{cases} 
	%	\frac{1}{H_{j_p}} & (i, j) \in h_p \\
	%	0 & \text{otherwise.} \\
	%\end{cases}$$
	%
	\begin{align}
		\label{eqn:gibbs1}
		P\left(h\left(Z_j^{(s+1)}\right) = p \mid \Phi, \bm{Z^{(s)}}\right) \propto
		\begin{cases} 
			w_{p}\times H_{j_p}  & p \leq P; \\
			n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\lambda}}{n_{12}(\bm{Z}) + \alpha_{\lambda}} &   p = P + 1 \\
		\end{cases}
	\end{align}
	\begin{align}
		\label{eqn:gibbs2}
		P\left(Z_j^{(s+1)} = i \mid h\left(Z_j^{(s+1)}\right) = p\right) = \begin{cases} 
			\frac{1}{H_{j_p}} & (i, j) \in h_p \\
			0 & \text{otherwise.} \\
		\end{cases}
	\end{align}

@@

	In Figure~\ref{fig:speed1}, \texttt{BRL} is faster than \texttt{fabl} for low sample sizes, but that \texttt{fabl} is significantly faster at handling larger data. In particular, run-time for \texttt{BRL} seems to grow quadratically (or linearly with the size of both $\bm{X}_1$ and $\bm{X}_2$) while run-time for \texttt{fabl} seems to grow linearly (in the size of $\bm{X}_2$).

>>>>

\begin{abstract}
Recently, researchers have developed Bayesian versions of the Fellegi Sunter model for record linkage. These have the crucial advantage of quantifying uncertainty from imperfect linkages.  However, current implementations of Bayesian Fellegi Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks.  We propose a variation on Bayesian Fellegi Sunter models that we call fast beta linkage, or \texttt{fabl}.  Specifically, we propose independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead. This also allows us to complete pairwise record comparisons over large datasets through parallel computing and to reduce memory costs through a new technique called storage efficient indexing. Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy.
\end{abstract}

@@

In what follows, Section \ref{sec:review-of_prior-work} reviews the work of \cite{fellegi_theory_1969} and \cite{sadinle_bayesian_2017}. Section \ref{sec:fast-beta-linkage}, proposes the \texttt{fabl} model, derives the Gibbs sampler for posterior inference, and provides the loss function used to calculate the Bayes estimate for the bipartite matching. Section \ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculatoins and the scale of linkage tasks amenable to \texttt{fabl}. Sections \ref{sec:simulations} and \ref{sec:case-studies} demonstrate the speed and accuracy of \texttt{fabl} through simulation studies and case studies of homicides from the El Salvadoran Civil War and the National Long Term Care Study.

@@

Importantly, this independence relaxes the one-to-one requirement from \texttt{BRL}; the Gibbs sampler under \texttt{fabl} does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\). To resolve many-to-one matchings and obtain a one-to-one estimate, we use the post processing procedure shown in Section \ref{bayes-estimate}.

@@

We initialize $\Phi$ from random draws from the prior distributions of $\bm{m}$ and $\bm{u}$, and initialize $\bm{Z}$ to reflect no matches across datasets, that is $\bm{Z} = (n_1 + 1, \ldots, n_1 + n_2)$. We then sample $\Phi$ from its full conditional:

$$\bm{m}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\alpha_{f1}(\bm{Z}), \ldots, \alpha_{fL_f}(\bm{Z})),$$
$$\bm{u}_f|\bm{Z}, \Gamma \sim \text{Dirichlet}(\beta_{f1}(\bm{Z}), \ldots, \beta_{fL_f}(\bm{Z})),$$
where $\alpha_{fl}(\bm{Z})= \alpha_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j = i)$, and $\beta_{fl}(\bm{Z})= \beta_{fl} + \sum_{i,j} I_{obs}(\gamma_{ij}^f)I(\gamma_{ij}^f = l) I(z_j \neq i)$.

Next, we sample $\bm{Z}$ componentwise from the full conditionals for each $Z_j$:
$$p\left(Z_j^{(s+1)}  = i| \Phi, \bm{Z^{(s)}}\right) \propto
\begin{cases} 
	w_{z_j, j},  & z_j \leq n_1; \\
	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}}, & z_j  = n_1 + j, \\
\end{cases},$$
where 
$$w_{ij} = \prod_{f=1}^{F}\prod_{l = 1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.$$
Derivations for these full conditionals can be found in Appendix \ref{app:derivations}.

@@

To increase the speed of the Gibbs sampler and expand the scale of the linkage tasks we can undertake, we propose hashing methods similar to those used by \cite{enamorado2019using} in the creation of \texttt{fastlink}, a fast and scalable implementation of the Fellegi Sunter method. The key insight is to recognize that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector. To make this more precise, let $h_1, \ldots, h_P$ denote the unique agreement patterns, and collect these unique patterns in the set $\mathcal{P} = \{h_1, \ldots, h_P\}$. Here,  $P = |\mathcal{P}|$ is the total number of unique agreement patterns. \(P\) is bounded above by \(\prod_{f=1}^F L_f\) which does not depend on \(n_1\) or \(n_2\). In this context, the integers $\{1, \ldots, P\}$ serve as hashed values that encode the same information as the orignial vectors themselves. Whenever possible, we conduct calculations over these \(P\) agreement patterns, instead of the typical \(n_1 \times n_2\) record pairs. Additionally, we store ``one-hot encodings'' of these patterns rather than the original Fellegi Sunter comparison vectors, to aid in vectorized computations. See Appendix \ref{app:ohe} for details.

@@

With SEI, however, we can reduce the comparison space, and its associated storage costs, while avoiding these drawbacks. Observe that all records $i \in \bm{X}_1$ that share the same agreement pattern with $j \in \bm{X}_2$ have the same Fellegi Sunter weight $w_{ij}$, and therefore the same probability when sampling \(Z_j\). Thus, we know that records $i \in r_{j_p}$  such that $H_{j_p}$ is large are very unlikely to be sampled consistently enough to be deemed a match through the Bayes estimate. We know this regardless of the form of the agreement pattern itself, or its associated probabilities. Therefore, rather than store all of these unlikely record labels, we choose to store only a small number \(S\) of them in a new nested listed $\mathcal{R}^{\text{SEI}}$. Rather than storing \(n_1 \times n_2\) record labels, SEI allows us to store at most \(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be. \textcolor{red}{In practice, we recommend $S=10$, as this reduces the number of stored indices for highly unlikely record pairings, but does is not likely to eliminate any of the indices for likely matches. Choosing $S$ too low, like $S=1$ or $S=2$, can concentrate undue mass on unlikely matches and distort linkage results.}Posterior calculations still attribute the appropriate weight to all records through the summary statistics in $\mathcal{H}$, and thus we can proceed with posterior inference through the memory reduced $\tilde{\Gamma}^{\text{SEI}} = \{\mathcal{P}, \mathcal{R}^{\text{SEI}}, \mathcal{H}\}$. For simplicity, we suppress ``SEI'' in future notation, and use $\tilde{\Gamma}$ without loss of generality.

@@

Instead, we use one-hot encodings of the agreement patterns $\mathcal{P}$ for more efficient calculations. Denote $H_p^m = \sum_{j=1}^{n_2} I\left((Z_j, j) \in h_p\right)$ to be the number of matching record pairs with agreement pattern $p$. The number of non-matching record pairs with agreement pattern $p$ is \(H_p^u = H_p - H_p^m\). Then, if $\bm{\alpha_0}$ and $\bm{\beta_0}$ are vectors of prior parameters for the $\bm{m}$ and $\bm{u}$ distributions respectively, the posterior update becomes
$$\bm{\alpha(Z)} = \bm{\alpha_0} + \sum_{p=1}^P H_p^m \times h_p,$$
$$\bm{\beta(Z)} = \bm{\beta_0} + \sum_{p=1}^P H_p^u \times h_p.$$
These constitute $\tilde{L}$ summations over $P$ quantities, where $\tilde{L} = \sum_{f=1}^F L_f$ is the total number of levels used across all features in the linkage task.

Although sampling \(Z_j\) from a the full conditional provided in Section \ref{gibbs_sampling} is conceptually straightforward, it becomes computationally burdensome when $n_1$ is large. This is because sampling a value from $n_1$ many options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_1$. To speed up computation, we break this sampling step into two. First, we calculate the Fellegi Sunter weight \(w_{p}\) associated with each unique pattern and sample the agreement pattern between \(j\) and its potential link. Second, we sample the record label uniformly among records associated with that agreement pattern for that particular \(j\in B\). More concretely, define \(h(Z_j)\) to be the agreement pattern between \(j\) and its potential match, and say \(h(Z_j) = h_{P+1}\) when \(Z_j = n_1 + j\). Then,

%$$P\left(h\left(Z_j^{(s+1)}\right) = p \mid \Phi, \bm{Z^{(s)}}\right) \propto
%\begin{cases} 
%	w_{p}\times H_{j_p}  & p \leq P; \\
%	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\lambda}}{n_{12}(\bm{Z}) + \alpha_{\lambda}} &   p = P + 1 \\
%\end{cases}$$
%$$P\left(Z_j^{(s+1)} = i \mid h\left(Z_j^{(s+1)}\right) = p\right) = \begin{cases} 
	%	\frac{1}{H_{j_p}} & (i, j) \in h_p \\
	%	0 & \text{otherwise.} \\
	%\end{cases}$$
	%
	\begin{align}
		\label{eqn:gibbs1}
		P\left(h\left(Z_j^{(s+1)}\right) = p \mid \Phi, \bm{Z^{(s)}}\right) \propto
		\begin{cases} 
			w_{p}\times H_{j_p},  & p \leq P; \\
			n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\lambda}}{n_{12}(\bm{Z}) + \alpha_{\lambda}}, &   p = P + 1; \\
		\end{cases}
	\end{align}
	\begin{align}
		\label{eqn:gibbs2}
		P\left(Z_j^{(s+1)} = i \mid h\left(Z_j^{(s+1)}\right) = p\right) = \begin{cases} 
			\frac{1}{H_{j_p}}, & (i, j) \in h_p; \\
			0, & \text{otherwise.} \\
		\end{cases}
	\end{align}

@@

	In Figure \ref{fig:speed1}, \texttt{BRL} is faster than \texttt{fabl} for low sample sizes, but that \texttt{fabl} is significantly faster at handling larger data. In particular, run-time for \texttt{BRL} seems to grow quadratically (or linearly with the size of both $\bm{X}_1$ and $\bm{X}_2$) while run-time for \texttt{fabl} seems to grow linearly (in the size of $\bm{X}_2$).
